### 2023-03-06 README.Rmd
init repo

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# simulating-meta-analysis

<!-- badges: start -->
<!-- badges: end -->