### 2023-03-07 README.Rmd
updating readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Understanding meta-analysis through data simulation, with applications to power analysis

<!-- badges: start -->
[<img alt="alt_text" src="https://img.shields.io/badge/OSF-https%3A%2F%2Fosf.io%2F54djn%2F-blue" />](https://osf.io/54djn/)
<!-- badges: end -->

This repository contains the code to reproduce the analysis of the paper **Understanding meta-analysis through data simulation, with applications to power analysis** by Gambarota Filippo and Gianmarco Altoè. The project can also be found on the Open Science Framework https://osf.io/54djn/.

## Repository

The project is organized as follows:

- `R`: contains the simulation functions used in the paper and supplementary materials
- `docs`: contains the `.Rmd` files to reproduce the paper and supplementary materials manuscripts
- `scripts`: contains the script used to produce complex objects and figures
  - `scripts\r-img.R`: create the paper figures. Figures that are not produced in this script using `ggplot2` are created with the open-source program [*Inkscape*](https://inkscape.org/) and contained in the `../img/` folders.
- `objects`: contains the result of objects created by the `scripts/` scripts.

The project is organized and structured as an R package, thus after opening the `simulating-meta-analysis.Rproj` file and installing the `devtools` package, calling `devtools::load_all()` will load each function within the `R/` folder.